---
title: "Tidy Tuesday 4"
description: |
  This post will explore Professor Lisa lendway's Garden Dataset 2021 with the objective of producing informative visualizations
author:
  - name: Tshering Wangchuk
    url: {}
date: 2021-10-25
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = FALSE)
```

### Loading libraries

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
library(googlesheets4)
library(lubridate)
library(ggthemes)
```

### Loading dataset - Lisa Lendway's 2021 Garden Data

```{r}
gs4_deauth() 

harvest_2021 <- read_sheet("https://docs.google.com/spreadsheets/d/1DU6dpxrbPGW2oJHQ6TG_xibazoGhvmRcqk2iPGOaRUI/edit?usp=sharing") %>% mutate(date = ymd(date))
```

### Visualizations 

<br> Since I come from South Asia and eat peppers quite frequently, I was curious to understand patterns for the different type of peppers within the dataset. I decided to create a visualization that would allow us to understand each variety in comparison to eachother. The graph I produced informs uf of the different harvest rate for each type of pepper. 

```{r}
harvest_2021 %>% filter(vegetable %in% c("peppers")) %>%
                 group_by(variety) %>% 
                 summarise(weight = n()) %>% 
                 ggplot(aes(x = reorder(variety, weight), y = weight, , fill = variety)) + 
                 geom_bar(stat = "identity") + coord_flip() + 
                 labs(title = "Harvest rate for different varieties of Pepper", 
                      y = "Number of times harvested", x = "Varieties of Vegetables",
                      caption = " Source: Tshering Wangchuk, Macalester College") + theme_clean() +
                      theme(legend.position = "none")
```



