---
title: "Tidy Tuesday 2"
description: |
  This post will explore how features / characteristics of music production affect their popularity on Spotify
author:
  - name: Tshering Wangchuk
    url: {}
date: 09-22-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading Library

```{r}
library(dplyr)
library(ggplot2)
library(ggthemes)
```

### Loading dataset

```{r}
audio_features <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-09-14/audio_features.csv')
```

### Data Visualization 

```{r}
 Audio <- audio_features %>% 
          filter(performer %in% c("Drake", "Beyonce", "Polo G", "Ariana Grande")) %>%
          group_by(performer) %>%
          select(spotify_track_popularity, danceability)

Audio %>% ggplot(aes(x = danceability, y = spotify_track_popularity)) + geom_smooth(method = lm) + 
          labs(title = "The Relationship Between Popularity and Danceability on Spotify",
                    x = "Danceability",
                    y = "Popularity",
                    caption = " Source: Tshering Wangchuk, Macalester College") + 
          theme_clean()
```
<br>
<br> The goal of using this dataset was to obersve characteristics / features of music that increased it's popularity on Spotify. I selected four artists that have a large following among the younger generation and the plot displays that the chances of a song being liked on spotify is positively correlated with danceability. 


